Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread

### 2018-02-08 paper.tex
comments

Adaptive introgression has three components: hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and selection against undesirable wild background.
To date, literature on crop-wild gene flow has focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
The history of natural and potentially adaptive introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
Preliminary results suggest a need to expand our conception of domestication to include the broadening of the genetic base of crops that occurred during post-domestication expansion through gene flow with newly encountered wild relatives.

@@

\item{Potato}

\gmj{I had saved these notes from an earlier version of the paper.  There is plenty of evidence of hybridization and introgression (both farmer-mediated and and natural), and some adaptive traits have been conferred, but I could not find direct evidence of specific cases of adaptive natural introgression from wild relatives to cultivated potatoes. This new paper fills that gap.}

Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 years ago, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}.
The northern members of the polyphyletic \emph{S. brevicaule} wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because \emph{s. tuberosum} has a polyphyletic origin.
Also, these northern members of the \emph{S. breviaule} complex are not clearly defined, and may in fact be one singular species (in which case the species name would be \emph{S. bukasovii}, \cite{spooner2005single}).

Although potatoes are usually propogated clonally, farmers also promote sexual hybridization at times to improve disease resistance and develop new cultivars \cite{quiros1992increase}.
Farmers continue to grow potatoes in close proximity to wild relatives, resulting in domesticate-weedy-wild hybrid complexes which promote introgressive hybridizatiion \cite{rabinowitz1990high, johns1987relationships, linder1987diversity}.
These complexes, combined with a diverse range of biotic and environmental selective pressures and local farming practices (human-mediated migration, isolated farmsteads in fertile valleys, clonal propogation, and intentional maintenance of a variety of landraces), have fostered expansion of genetic diversity within potatoes subsequent to domestication \cite{brush1995potato}.
However, as farmers tend to abandon fields after being used for potato cultivation, it is less likely that hybrids have an opportunity to form stable populations for maintained introgressive gene flow into the domesticated potato gene pool \cite{brush1995potato}.
Also, natural selection may be subverted by farmer preference and artificial selection \cite{brush1981dynamics}.

Introgressive hybridization is widespread in potatoes \cite{grun1990evolution}.
The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}.
Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}.
Cultivated diploid potatoes have been shown introgressed with genes from \emph{S. sparsipilum} \cite{rabinowitz1990high} and \emph{S. megistacrolobum} \cite{johns1987relationships, huaman1980solanum}.
Drought- and frost-resistance genes introgressed from wild relatives expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}.
Resistance genes have been experimentally introgressed from wild poptato relatives (\emph{S. tuberosum subsp. andigena}, \cite{van1999tight}, \emph{S. bulbocastanum}, \cite{van2003ancient}) into cultivated potato, although this process is hindered by compatability issues including ploidy levels and Endosperm Balance Number \cite{johnston1980significance}.

@@

\begin{table}
\rowcolors{2}{white}{gray!25}
\begin{center}
\caption{List and brief description of recently developed methods and examples of empirical studies employing these methods.} \label{tab:tools}
\begin{tabular}{llll}
\\\toprule  
\rowcolor{white}
{\bf methods}	& {\bf data type } &	{\bf reference} &  {\bf empirical studies } \\ \midrule

@@

\subsection*{When was a crop domesticated?}
Estimates of the timing of initial domestication are often based on levels of sequence divergence between a crop and populations of its presumed progenitor (\emph{e.g.}, \cite{matsuoka2002single, molina2011molecular}).
In highly introgressed domesticates, these estimates will be based on comparison of either crop or introgressant haplotypes to those of the progenitor.
In such cases, divergence is a mixture of time since domestication and time since split of the progenitor and the introgressing species.
This phenomenon, in combination with divergence of samples from true ancestral populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may explain discrepancies between domestication dates based on genetic and archaeological data.
More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression.

\subsection*{How was genome-wide diversity impacted by a domestication bottleneck?}
Measurement of the strength of the initial domestication bottleneck may also be impacted by adaptive introgression during the spread of crops.
Crop wild relatives have distinct demographies when compared to domesticates and may therefore have contrasting effective population sizes ($N_e$).
The influence of wild relative introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of gene flow, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
For example, substantial introgression at neutral loci from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the domestication bottleneck.
\gmj{Does Li's research fit in well here? Should we add her citation?}

@@

After hybrization events that lead to introgression, how long might the detectable genomic signals of introgression persist?
%\gmj{Depends on number of generations, effective population size, mutation rate, recombination rate, patterns and strength of selection.}
%\gmj{If an introgression from the progenitor sweeps across the cultivated species, how is it distinguished from a genomic region that is identical by descent and wasn't lost in the domestication process?}
%Does one see different architectures of introgression across different regions?
%Have introgressed alleles been constrained to the population which initially recieved the introgression, or have those alleles spread to the rest of of the crop species?
Introgressed regions are easier to detect when there has been limited recombination to break them apart.
Therefore, introgressions are easiest to detect when they are either recent (few generations means few recombination events) or involve structural variation (which diminishes recombination rate).
Because recombination progressively breaks apart LD in introgressed regions, measurements of LD can be used to date introgression events (as in \cite{Poets2015}).

>>>>

Adaptive introgression has three components: hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and selection against undesirable wild background.
To date, literature on crop-wild gene flow has focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
The history of natural and potentially adaptive introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
Preliminary results suggest a need to expand our conception of domestication to include the broadening of the genetic base of crops that occurred during post-domestication expansion through gene flow with newly encountered wild relatives.
\lwang{I think these paragraphs read pretty smooth. I pondered at the last sentence of this paragraph. 
Suggestions: Preliminary results suggest a need to expand our conception of genetic bases of domestication in crops that inhabited new territory through adaptive gene flow from newly encountered wild relatives during post-domestication expansion.}

@@

@@

\begin{table}

\rowcolors{2}{white}{gray!25}
\begin{center}
\caption{List and brief description of recently developed methods and examples of empirical studies employing these methods.} \label{tab:tools}
\begin{tabular}{0.9\textwidth}{llll}
\\\toprule  
\rowcolor{white}
{\bf methods}	& {\bf data type } &	{\bf reference} &  {\bf empirical studies } \\ \midrule

@@

\subsection*{When was a crop domesticated?}
Estimates of the timing of initial domestication are often based on levels of sequence divergence between a crop and populations of its presumed progenitor (\emph{e.g.}, \cite{matsuoka2002single, molina2011molecular}).
In highly introgressed domesticates, these estimates will be based on comparison of either crop or introgressant haplotypes to those of the progenitor.
In such cases, divergence is a mixture of time since domestication and time since split of the progenitor and the introgressing species.
This phenomenon, in combination with divergence of samples from true ancestral populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may explain discrepancies between domestication dates based on genetic and archaeological data.
\lwang{maybe some citations for a protracted domestication time: African rice NG paper, our paper, grape PNAS paper}
More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression.
\lwang{excluding the introgression regions might cause some inference method not working properly, such as MSMC, as it would reduce heterozygosity.
Suggestions: "More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression and / or developing new statistic tools which taking potential introgression into account."}

\subsection*{How was genome-wide diversity impacted by a domestication bottleneck?}
\lwang{I suggest to delete "impacted by a domestication bottleneck" because that might confuse readers. Adaptive introgression could affect genome-wide diversity, owing to that introgressing haplotype have a higher historical Ne and domesticates have a lower historical Ne due to bottleneck. Adding "impacted by a domestication bottleneck" seems making the message not direct and not complete.}
Measurement of the strength of the initial domestication bottleneck may also be impacted by adaptive introgression during the spread of crops.
Crop wild relatives have distinct demographies when compared to domesticates and may therefore have contrasting effective population sizes ($N_e$).
The influence of wild relative introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of gene flow, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
For example, substantial introgression at neutral loci from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the domestication bottleneck.
\gmj{Does Li's research fit in well here? Should we add her citation?}

@@

After hybrization events that lead to introgression, how long might the detectable genomic signals of introgression persist?
%\gmj{Depends on number of generations, effective population size, mutation rate, recombination rate, patterns and strength of selection.}
%\gmj{If an introgression from the progenitor sweeps across the cultivated species, how is it distinguished from a genomic region that is identical by descent and wasn't lost in the domestication process?}
%Does one see different architectures of introgression across different regions?
%Have introgressed alleles been constrained to the population which initially recieved the introgression, or have those alleles spread to the rest of of the crop species?
Introgressed regions are easier to detect when there has been limited recombination to break them apart.
Therefore, introgressions are easiest to detect when they are either recent (few generations means few recombination events) or involve structural variation (which diminishes recombination rate).
\lwang{involve --> involving}
Because recombination progressively breaks apart LD in introgressed regions, measurements of LD can be used to date introgression events (as in \cite{Poets2015}).
